---
title: "Internally Displaced Persons Globally"
subtitle: "Due to Conflict/Violence and Disasters"
author: "Adam Yeazel"
institute: "SIS 750: Data Analysis"
date: "04/07/2023"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["uwm", "uwm-fonts"]
    nature:
      countdown: 60000
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(patchwork)
library(leaflet)
library(ggthemes)
library(lattice)

setwd("/Users/adamyeazel/Desktop/American University/SIS 750 Data Analysis/Problem Sets/Presentation 2")

# reading in data
idp = readxl::read_excel('IDMC_Internal_Displacement_Conflict-Violence_Disasters_2008_2021.xlsx')
```

# Data

The data for this presentation comes from the [Internal Displacement Monitoring Centre](https://www.internal-displacement.org/database/displacement-data) one of the leading sources of data and analysis on internal displacement and part of the Norwegian Refugee Council (NRC). 

--

- The data contains 1,965 observations of 11 variables; observations are years between 2008-2021 grouped by country

--

- The data includes displacements for a given year as well as stock numbers of IDPs who remain displaced from previous years

--

```{r data, echo = FALSE, warning=FALSE, message=FALSE}
# summary table for both conflict and disasters

# cleaning the first row of descriptions of variables 
df = idp[-1,]

# changing data types to be numeric
df$`Conflict Internal Displacements` <- as.numeric(df$`Conflict Internal Displacements`)
df$`Conflict Stock Displacement` <- as.numeric(df$`Conflict Stock Displacement`)
df$`Disaster Internal Displacements` <- as.numeric(df$`Disaster Internal Displacements`)
df$`Disaster Stock Displacement` <- as.numeric(df$`Disaster Stock Displacement`)

# making a summary table of conflict IDPs 
df %>%
  summarize(
    Min = min(`Conflict Internal Displacements`, na.rm = TRUE),
    Mean = mean(`Conflict Internal Displacements`, na.rm = TRUE),
    Median = median(`Conflict Internal Displacements`, na.rm = TRUE),
    Max = max(`Conflict Internal Displacements`, na.rm = TRUE),
    SD = sd(`Conflict Internal Displacements`, na.rm = TRUE)
  ) %>% 
  knitr::kable(format = 'html', digits = 1L,
               caption = "Conflict Internal Displacements (per Year)")
```
<br/>
```{r table 2, echo=FALSE, warning=FALSE, message=FALSE}
# summary table of disaster IDPs
df %>%
  summarize(
    Min = min(`Disaster Internal Displacements`, na.rm = TRUE),
    Mean = mean(`Disaster Internal Displacements`, na.rm = TRUE),
    Median = median(`Disaster Internal Displacements`, na.rm = TRUE),
    Max = max(`Disaster Internal Displacements`, na.rm = TRUE),
    SD = sd(`Disaster Internal Displacements`, na.rm = TRUE)
  ) %>% 
  knitr::kable(format = 'html', digits = 1L,
               caption = "Disaster Internal Displacements (per Year)")
```

---

# Data

- The data is extremely right skewed, so it has been log transformed to facilitate analysis

```{r summary viz, echo=FALSE, fig.height=6, fig.width=12, message=FALSE, warning=FALSE, dpi=300}
# creating log histogram for conflict IDPs
plot1 = ggplot(data = df, aes(log(`Conflict Internal Displacements`))) + 
  geom_histogram(color = "white", fill = "#c5050c") + 
  scale_y_continuous(limits = c(0, 150)) + 
  labs(x = "Log Conflict IDPs", y = "Count", main = "Log Conflict IDPs") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 17))

# creating log histogram for disaster IDPs
plot2 = ggplot(data = df, aes(log(`Disaster Internal Displacements`))) + 
  geom_histogram(color = "white", fill = "#c5050c") + 
  scale_y_continuous(limits = c(0, 150)) + 
  labs(x = "Log Disaster IDPs", y = "Count ", main = "Log Disaster IDPs") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 17))

plot1 + plot2 + 
  plot_annotation(
    title = "Log Conflict IDPs vs. Log Disaster IDPs",
    caption = "Data up to 2021",
    theme = theme(plot.title = element_text(size = 25)) + 
    theme(plot.title = element_text(hjust = 0.5, face = 'bold')) +
    theme(axis.text.x = element_text(size = 20),
            axis.text.y = element_text(size = 20)))
```

---

class: inverse, middle, center

# Exploring the Data

---

# Case Studies due to Conflict vs. Disaster

```{r case studies, warning=FALSE, message=FALSE, echo=FALSE, fig.height=8, fig.width=12, dpi=300}
# forcing non-scientific notation
options(scipen = 90)

# graphing IDPs in Sudan over time 
sudan = df %>% 
  filter(Name == "Sudan")
afghanistan = df %>% 
  filter(Name == "Afghanistan")
yemen = df %>% 
  filter(Name == "Yemen")

# log conflict displacements over time
c1 = ggplot(data = sudan, aes(Year, `Conflict Internal Displacements`, group = 1)) +
  geom_line(aes(color = "Sudan"), lwd = 2) + 
  geom_line(data = afghanistan, aes(color = "Afghanistan"), lwd = 2) +
  geom_line(data = yemen, aes(color = "Yemen"), lwd = 2) + 
  labs(y = "Log Conflict IDP", color = "Country", lwd = 2) + 
  scale_y_continuous(trans = "log10") + 
  annotation_logticks(sides = "l") + 
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        axis.text.x = element_text(angle = -30, vjust = 1, hjust = 0, size = 15),
        axis.text.y = element_text(size = 15),
        axis.title = element_text(size = 17)) + 
  theme_minimal() + 
  plot_annotation(
    caption = "Data up to 2021",
    theme = theme(plot.title = element_text(size = 30)) + 
    theme(plot.title = element_text(hjust = 0.5, face = 'bold')) +
    theme(axis.text.x = element_text(size = 25),
            axis.text.y = element_text(size = 25))) + 
  theme(legend.text = element_text(size = 15), legend.title = element_text(size = 15))

# log disaster displacements over time 
c2 = ggplot(data = sudan, aes(Year, `Disaster Internal Displacements`, group = 1)) +
  geom_line(aes(color = "Sudan"), lwd = 1, linetype = 5) + 
  geom_line(data = afghanistan, aes(color = "Afghanistan"), lwd = 1, linetype = 5) +
  geom_line(data = yemen, aes(color = "Yemen"), lwd = 1, linetype = 5) + 
  scale_y_continuous(trans = "log10") + 
  annotation_logticks(sides = "l") + 
  labs(y = "Log Disaster IDP", color = "Country", lwd = 2) + 
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        axis.text.x = element_text(angle = -30, vjust = 1, hjust = 0, size = 15),
        axis.text.y = element_text(size = 15),
        axis.title = element_text(size = 17))+ 
  theme_minimal() +
  plot_annotation(
    caption = "Data up to 2021",
    theme = theme(plot.title = element_text(size = 30)) + 
    theme(plot.title = element_text(hjust = 0.5, face = 'bold')) +
    theme(axis.text.x = element_text(size = 25),
            axis.text.y = element_text(size = 25))) + 
  theme(legend.text = element_text(size = 15), legend.title = element_text(size = 15))

c1 / c2 + 
  plot_annotation(title = "Conflict vs. Disaster IDPs With Case Studies",
                  caption = "Data up to 2021",
    theme = theme(plot.title = element_text(size = 30)) + 
    theme(plot.title = element_text(hjust = 0.5, face = 'bold')) +
    theme(axis.text.x = element_text(size = 40),
            axis.text.y = element_text(size = 40))) + 
  theme(legend.text = element_text(size = 40))
```
